---
title: "EDA with group_by"
author: "Jingyi Yao"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

```


```{r}
library(tidyverse)
library(ggridges)
```


### `lubridate` package inside `tidverse`
### `floor_date` : round the date to the **first** day in the month it belongs to
```{r,message=FALSE}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())

weather_df
```


## 1. `group_by()`
```{r}
weather_df %>%
  group_by(name, month)
```




## 2. counting things

### `n()` 
### `summarize(new column = n())` count in group_by() and summarize()
### store the number of each group in the new column
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(n_obs = n())
```

```{r}
weather_df %>% 
  group_by(name) %>% 
  summarize(
    n_obs=n()  # n() can calculate the number
  )

```




### `count(column,name="new column")`
### count the column and store the result in the new column
### do not need to group_by and summarize
```{r}
weather_df %>%
  count(month, name = "n_obs")
```



### table is not recommended
```{r}
weather_df %>%
  pull(month) %>% 
  table
```



### multiple summaries -- count many items in summarize()
### `n_distinct(something)` count the unique items in each group
```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```



## 3. 2 by 2 table

### use `case_when()` to input (condition -- value) results
```{r}
weather_df %>% 
  mutate(
    cold = case_when(
      tmax <  5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE      ~ ""           # last line in case_when to remind you if there is a problem
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```


### `janitor::tabyl(column1,column2)` to make 2 by 2 table 
### use janitor::tabyl() and we do not need group_by() + sumarize()
```{r}
weather_df %>% 
  mutate(cold = case_when(
    tmax <  5 ~ "cold",
    tmax >= 5 ~ "not_cold",
    TRUE     ~ ""
  )) %>% 
  filter(name != "Waikiki_HA") %>% 
  janitor::tabyl(name, cold)
```



## 4. General Summaries

```{r}
weather_df %>%
  group_by(month) %>%
  summarize(
    mean_tmax = mean(tmax),
    mean_prec = mean(prcp, na.rm = TRUE),
    median_tmax = median(tmax),
    sd_tmax = sd(tmax))

```


```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(
    mean_tmax = mean(tmax),
    median_tmax = median(tmax))

```

### use `summarize(across(column1:column10, function))` function is in across
### if you want to apply one function to many many columns, use across in summarize()
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(across(tmin:prcp, mean))
```


## geom_point() + geom_line() or geom_path() : show the lines and the points together
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax,na.rm=TRUE)) %>%
  ggplot(aes(x = month, y = mean_tmax, color = name)) + 
    geom_point() + geom_path() + 
    theme(legend.position = "bottom")

```


### `knitr::kable(digits = 1)` appended as the last step
### very useful when we print out a tibble or dataframe
### print a table format dataframe with a better look
### what does digits mean ?
```{r}
weather_df %>%
  group_by(name, month) %>%
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax) %>% 
  knitr::kable(digits = 2)
```




## 5. Grouped `mutate`

### not using `summarize()` after `group_by`
### use `mutate()` after `group_by`
### summarize() only shows the results of the summarize calculation
### mutate() shows the previous dataframe and the new columns added
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(mean_tmax = mean(tmax,na.rm=TRUE),
         centered_tmax =tmax - mean_tmax) %>%
  ggplot(aes(x = date, y = centered_tmax, color = name)) + 
    geom_point() + geom_line() + 
    theme(legend.position = "bottom")

```




## 6. lagged observations

### `lag(data,n = rows)` and `lead(data,n = rows)` 
### the parameter `n` specifies how many places(rows) we want to lag or lead
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(
    yesterday_tmax=lag(tmax),
    tmax_change = tmax - yesterday_tmax
  ) %>% 
  summarize(
    sd_tmax_change = sd(tmax_change,na.rm=TRUE)
  )

```




## 7. Window functions -- ranking

### `row_number` rank in ascending order and rank the same number with different orders
### `min_rank` also ranks in ascending order but gives the same number the same order
### `dense_rank` ranks in ascending order and gives the same number the same order but does not rank the next one with a plus 1
```{r}
x <- c(5, 1, 3, 2, 2,8,6,6, NA)
row_number(x)
min_rank(x)
dense_rank(x)

```

### `min_rank(column)`
```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(tmax_rank = min_rank(tmax))

```

```{r}
weather_df %>%
  group_by(name, month) %>%
  mutate(tmax_rank = min_rank(tmax)) %>% 
  filter(min_rank(tmax) < 4) %>% 
  arrange(name,month,tmax_rank)

```


```{r}
weather_df %>%
  group_by(name, month) %>%
  filter(min_rank(desc(tmax)) < 4)

```

### the difference between the original data and the lagged data
```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax))

```


```{r}
weather_df %>%
  group_by(name) %>%
  mutate(temp_change = tmax - lag(tmax)) %>%
  summarize(
    temp_change_sd = sd(temp_change, na.rm = TRUE),
    temp_change_max = max(temp_change, na.rm = TRUE))

```

