---
title: "ggplot2 part I"
author: "Jingyi Yao"
date: "`r Sys.Date()`"
output: html_document
---
### global settings of the layout format
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

```

## load `tidyverse` and `ggridges`
```{r}
library(tidyverse)
library(ggridges)
```



#### working with `NOAA weather` data
#### downloaded using `rnoaa::meteo_pull_monitors` function 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```



## Basic scatterplot
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()

```

### `drop_na()` to eliminate the rows with `NA`
```{r}
weather_df %>% drop_na() %>% ggplot(aes(x = tmin, y = tmax)) + geom_point()

# drop_na

```


```{r}
weather_scatterplot=
  weather_df %>% 
  drop_na() %>% 
  ggplot(aes(x = tmin, y = tmax)) + geom_point() 

# already claimed the weather_df at the front, do not claim it in ggplot()

```


## plotting more fancy

### color the points in `geom_point(aes(color=column))`
### add a smooth line in `geom_smooth()`
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) +  # aes in geom_point()
  geom_smooth()                    # does it looks like a linear association?

```



### the `ggplot(aes())` applies to the whole plot
### the `geom_point(aes())` only to the points
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax,color = name)) + # aes color in ggplot too
  geom_point(aes(color = name)) +  # aes in geom_point()
  geom_smooth()                    

```


### set the alpha of points in `geom_point(aes(alpha=0.3))` the degree of transparency
### set the smooth vurve with `se=FALSE` whether shows the CI around the curve
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax,color=name)) + 
  geom_point(aes(color = name,alpha=.3)) +
  geom_smooth(se=FALSE)                    

```


### delete the points only shows the smooth lines by deleting `geom_point()`
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax,color=name)) + 
  geom_smooth()                    

```


### facet using `face_grid(y column ~ x column)` line the plots in columns
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax,color=name)) + 
  geom_point(aes(color = name,alpha=.3)) +
  geom_smooth(se=FALSE) +
  facet_grid(.~name)

```


### change the facet direction line the plots in rows
```{r}
ggplot(weather_df, aes(x = tmin, y = tmax,color=name)) + 
  geom_point(aes(color = name,alpha=.3)) +
  geom_smooth(se=FALSE) +
  facet_grid(name ~ .)

```


### change the size of points in `geom_points(aes(size=column))`
### the size depends on the value of data in the specified column
```{r}
weather_df %>% 
  ggplot(aes(x=date,y=tmax,color=name))+
  geom_point(alpha=.3,aes(size = prcp))+
  geom_smooth(se= FALSE)+
  facet_grid(.~name)


```


### only shows the fitted lines
```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 

```


### `geom_hex()`show the points in **hexagons shapes**
```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```


## Univariate Plots -- histograms, barplots,violins

## histograms

### `geom_histogram` to make a histogram
```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()

```

### `ggplot(aes(fill=column))` setting the filling color for the histogram
```{r}
ggplot(weather_df, aes(x = tmax,fill=name)) + 
  geom_histogram()

```

### `geom_histogram(position = "dodge")` put the bars together
### `geom_histogram(binwidth= 1)` set the bin width
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)

```


## Density plots
### `geom_density(alpha,adjust,color)` to make density plot
### `alpha` is the degree of transparency of the filling
### `color` is the color of the fillin area
```{r}
ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")

```


## Boxplots
```{r}
ggplot(weather_df, aes(x = name, y = tmax,fill=name)) + geom_boxplot()
```

## Violin Plots

### `+ stat_summary(fun= ,color= )` add summary stat to the violin plot
### `fun = "median"` shows the median point in the violin plot
### `color =  "blue"` is the color or the median point, the color is in the quotation mark
```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")

```

## Ridge Plots

### ridge plot is the stacked density plot without overlapping
### `geom_density_ridgdes` remember the s in ridges
```{r}
ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)

```


## Saving and Embedding plots

### first, let's make a plot

```{r}
weather_scatterplot =
  weather_df %>% 
  ggplot(aes(x=date,y=tmax,color=name))+
  geom_point(aes(size=prcp),alpha=.3)+
  geom_smooth(se=FALSE)+
  facet_grid(.~name)

weather_scatterplot
```

### next, save the scatterplot
### set the file name, and size  in `ggsave()`
### `plot = previous plot name`
### `width = , height = `
```{r}
ggsave(file="weather_scatter_plot.pdf",plot=weather_scatterplot,width=8,height=5)
```


