---
title: "ggplot2 part 2"
author: "Jingyi Yao"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

```

## load `tidyverse` and `ggridges`
```{r,warning=TRUE,message=FALSE}
library(tidyverse)
library(ggridges)
```

### load the weather dataset
```{r,message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```


## scatterplot -- but better this time

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,color=name)) +  # color = name
  geom_point(alpha = .5)

```


## Labels

### `+ labs(x=" ",y=" ",caption=" ")` labs with an s and the titles are in quotation marks
### title=, x=,y=
### `caption` is an annotation to the plot
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )
```



## Scales -- change the axis and color

### customize the x-axis labels using `scale_x_continuous(breaks=,labels=)`
### `breaks = vector(where to mark the data)` numeric vector showing the location of tick marks
### `labels = vector(caption)` a character vector of the marks
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบ C", "0", "15"))
```



### customize the y-axis values using `scale_y_continuous(trans=,position=)`
### `sqrt` the values in y
### `trans = mathematical transformation function`
### `position = left or right` where to put the labels
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบC", "0", "15"),
    limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
```


### change the scale color using `hue`
### `scale_color_hue(name=column,h= the color range)` the color range is an **interval**
### `name` is the column that the colors are assigned to
### `h` is a numeric vector to store the color range
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(name = "Location", h = c(100,400))
```


### use the color map in the `viridis` package
### specify the color range is **discrete** or **continuous**
### `viridis::scale_color_viridis(name = "Location",discrete = TRUE)`
### `name=` is still the column
### `discrete = T OR F` means whether the column is discrete or continuous
```{r}
ggp_temp_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_temp_plot
```



## Themes -- +theme()

### `+theme(legend.position= "bottom")` change the legend position and the postion is a character
### the position can be `"none"` which means removing the legend
### append the `theme(legend.position=" ")` to a plot
```{r}
ggp_temp_plot + 
  theme(legend.position = "bottom") # a character in quotation marks
```

### change the background color into `bw` or `classic` or `excel` or `minimal`

### the order of `theme_bw` and `theme()` cannot be changed
### `theme()` should be the last line
### white
```{r}
ggp_temp_plot + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

### excel
### grey with lines
```{r}
ggp_temp_plot + 
  ggthemes::theme_excel() + 
  theme(legend.position = "bottom")
```

### classic
### blank
```{r}
ggp_temp_plot + 
  theme_classic() + 
  theme(legend.position = "bottom")
```

### minimal
### white
```{r}
ggp_temp_plot + 
  theme_minimal() + 
  theme(legend.position = "bottom")
```

## Setting Option
```{r}

```


## Data in `geom()`

### `geom_line()` is the line that connects the points in order
```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```


## Patchwork (pacakge)

### using `facet` to line **related** graphs in one plot
### using `patchwork` package to place **different** graphs in one plot

#### plot separate plots first
```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

tmax_tmin_p

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

prcp_dens_p

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

tmax_date_p

```

### load the `patchwork` package
### use `+`, `/`,`()` to set the layout
```{r}
library(patchwork)
tmax_tmin_p + prcp_dens_p  # side by side

tmax_tmin_p / prcp_dens_p  # stacked

(tmax_tmin_p + prcp_dens_p) / tmax_date_p   # first side by side and then stacked

```



## Data Manipulation

### `forcats::fct_relevel()` the violin plot
### relevel by the order or name and the character vector specified
```{r}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

### `reorder()` the violin plot
```{r}
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```



### `relevel()` the boxplot
```{r}
weather_df %>% 
 mutate(name = fct_relevel(name,"Waikiki_HA")) %>%
 ggplot(aes(x=name,y=tmax))+geom_boxplot()
  
```

### `reorder()` the boxplot
```{r}
weather_df %>% 
 mutate(name = fct_reorder(name,tmax)) %>%
 ggplot(aes(x=name,y=tmax))+geom_boxplot()
  

```

### shows the distribution of bdi in each visit -- show the 4 boxplots side by side
### add *visit* as a variable in the dataset
### x=visit,y=bdi
### arrange the order in one way
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```

### arrange the order in another way -- `relevel()``
```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = fct_relevel(visit,"bl")   # use the fct_relevel()
  )

ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, pn_day, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```






